Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin

### 2017-10-23 paper/manuscript.tex
updated methods about trees

\subsection*{Construction of HA phylogenetic tree}
We downloaded one HA sequence per subtype from the NIAID Influenza Research Database (IRD) \comment{cite}, built the phylogenetic tree using RaxML \comment{cite}, and visualized the tree using FigTree \comment{cite}. The HA sequences we downloaded were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).

For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 in~\cite{dreyfus2013structure}. For subtypes not tested, we used the identity from the corresponding strain listed above.

We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}. Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; no antibodies have been tested against H17 and H18.

>>>>

\subsection*{Inference of HA phylogenetic tree}
To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{?}, inferred the phylogenetic tree using RaxML~\citep{?} with a \comment{what model of substitution?}, and visualized the tree using FigTree (\comment{url}). 
The HA sequences used were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize. 
Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs~\citep{caton1982antigenic}, we assume that they do not bind other subtypes.

For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by \citet{dreyfus2013structure}. 
For subtypes not tested, the amino-acid identity reported is that in the strain listed above.
